### 2011-04-14 Spatial_adaptation/local-adaptation.tex
Undo revision 3036 by [[Special:Contributions/Peter|Peter]] ([[User talk:Peter|talk]])

Here is one idea for computing the migration rate.
Suppose that there is a patch at the origin with $N$ individuals,
and we want to find the ``migration rate'' to another patch at distance $r$.
Each generation, offspring are divided into two classes:
those living in the good patch (i.e.\ replacing some of those $N$ individuals)
and new migrants, that might eventually make it to the other patch.
Suppose that in $dy$ there are approximately $m(y) dy$ of these latter.
Let $p(y,r)$ be the probability that the branching process begun at $y$ survives
{\em in the selective environment with all patches except the one at $r$ erased}.
If we can assume that the two are independent,
we can approximate the probability that the new patch is colonized by ``new migrants'' in a certain generation by
$\int m(y) p(y,r) dy$.

Here is another.  
\[
    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[Z_t(x) \one_{Z_t(A)=0}] ,
\]
so if $\E[Z_t(x) | {Z_t(A)=0}] \approx \E[Z_t(x)]$ 
then a good approximation for the rate may be
\[
    \int (1+s(x)) \kappa(x,A) \E[Z_t(x)] dx .
\]
{\bf Or,} let $\hat Z_t$ be the process with $s(A)=-\infty$.
Then $\E[ \hat Z_t(A) ] \approx \E[Z_t(A) \one_{Z_t(A)=0}]$,
so
\[
    \partial_t \P\{ Z_t(A) = 0 \} \approx - \sum_x (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] \P\{ Z_t(A) = 0\},
\]
and the approximation is
\[
    \int (1+s(x)) \kappa(x,A) \E[\hat Z_t(x)] dx ,
\]
which could by found as in Slatkin.

What about the stable distribution?
Suppose that mutants have selective (dis-)advantage $s(x)$ at location $x$,
that the symmetric dispersal kernel is $\kappa(\cdot)$, and that the mean-squared dispersal distance is $\sum_y y^2 \kappa(y) = \sigma^2$.
Then in the large population limit, if $\phi(t,x)$ is the proportion mutants at location $x$ at time $t$,
if $s$ and $\sigma^2$ are of the same (small) order,
(see Peter's slides from Berkeley probability seminar)
\begin{align}
    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
            + \sum_y \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) (\phi(t,x)-\phi(t,y))  \\
        &\approx s(x) \phi(t,x) (1-\phi(t,x)) + \frac{1}{2} \sigma^2 \partial_x^2 \phi(t,x) .
\end{align}
...

{\bf Note:} In some applications, 
it may be better to begin with the migration rates rather than using the theory above to estimate them.

>>>>

Suppose that mutants have selective (dis-)advantage $s(x)$ at location $x$,
that the symmetric dispersal kernel is $\kappa(\cdot)$, and that the mean-squared dispersal distance is $\sum_y y^2 \kappa(y) = \sigma^2$.
Then in the large population limit, if $\phi(t,x)$ is the proportion mutants at location $x$ at time $t$,
if $s$ and $\sigma^2$ are of the same (small) order,
(see Peter's slides from Berkeley probability seminar)
\begin{align}
    \partial_t \phi(t,x) &\approx \sum_y s(x) \kappa(y-x) (1-\one_{x=y}) \phi(t,y)(1-\phi(t,x)) 
            + \sum_y \kappa(y-x) (1-\frac{1}{N}\one_{x=y}) (\phi(t,x)-\phi(t,y))  \\
        &\approx s(x) \phi(t,x) (1-\phi(t,x)) + \frac{1}{2} \sigma^2 \partial_x^2 \phi(t,x) .
\end{align}
